---
title: "Global Terrorism Database"
output: 
  flexdashboard::flex_dashboard:
    theme: united
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(ggplot2)
library(leaflet)
library(shinyWidgets)
library(plotly)

#Load Packages
library(tidyverse)
```

```{r set working directory, echo = FALSE}
#Set working directory

setwd("C:/Users/Eli/Desktop/Global_Terrorism_Dataset")
```

```{r load data}
gtd <- read.csv("csv_gtd_70to95.csv", header = TRUE)
```

```{r convert to tibble}
gtdtib <- as_tibble(gtd)
```


Overview
===========================================================

Column {data-width=200 .sidebar}
-----------------------------------------------------------------------

### Global Terrorist Incidents By Region and Year

```{r}
selectInput("region", 
            "Select a Region of the World",
            choices = levels(gtdtib$region_txt))

sliderInput("year", 
            label = "Select a Year Range:",
            min = 1970,
            max = 1995,
            value = c(1970, 1995),
            step = 1,
            dragRange = TRUE)

gtd_filter1 <- reactive({
  gtdtib %>%
    filter(region_txt == input$region,
           iyear >= input$year[1] & iyear <= input$year[2]) %>%
    mutate(total = n())

})

```



Column {data-width=450}
-----------------------------------------------------------------------

### Map of Incidents Around the World

```{r}
renderLeaflet({
  
  gtd_filter1() %>%
    group_by(eventid, latitude, longitude) %>%
    count() %>%
    leaflet() %>%
    addTiles() %>%
    addCircles(radius = ~n)
  
})

```

Column {data-width=350}
-----------------------------------------------------------------------

### Number of Terrorist Incidents

```{r}
renderValueBox({
  valueBox(prettyNum(gtd_filter1() %>%
                       nrow(), big.mark = ','))
})
```

### Year Vs. Number of Casualties

```{r}


renderPlot({
  ggplot(data = gtd_filter1(), aes(x = iyear, y = nkill)) +
    geom_jitter(width = 0.3, height = 0, alpha = 0.3) +
    xlab("Year") +
    ylab("Number of Casualties")
  
})
```

Africa {data-navmenu="Middle East and Africa"} 
=========================================================


Column {data-width=250 .sidebar}
-----------------------------------------------------------------------

### Global Terrorist Incidents By Region and Year

```{r}

gtdtib_af <- filter(gtdtib, region_txt == "Sub-Saharan Africa")


pickerInput("countryaf", 
            "Select a Country In Sub-Saharan Africa",
            choices = levels(factor(gtdtib_af$country_txt)),
            selected = levels(factor(gtdtib_af$country_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

sliderInput("yearaf", 
            label = "Select a Year Range:",
            min = 1970,
            max = 1995,
            value = c(1970, 1995),
            step = 1,
            dragRange = TRUE)

pickerInput("attacktypeaf", 
            "Select a Type of Terrorist Attack",
            choices = levels(factor(gtdtib_af$attacktype1_txt)),
            selected = levels(factor(gtdtib_af$attacktype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

pickerInput("targettypeaf", 
            "Select Target Type",
            choices = levels(factor(gtdtib_af$targtype1_txt)),
            selected = levels(factor(gtdtib_af$targtype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

gtd_filter_af <- reactive({
  
  gtdtib_af %>%
    filter(country_txt %in% input$countryaf,
           iyear >= input$yearaf[1] & iyear <= input$yearaf[2],
           attacktype1_txt %in% input$attacktypeaf, 
           targtype1_txt %in% input$targettypeaf)
    })

```

Column {data-width=450}
-----------------------------------------------------------------------

### Map of Incidents Around the World

```{r}
renderLeaflet({
  
  gtd_filter_af() %>%
    group_by(eventid, latitude, longitude) %>%
    count() %>%
    leaflet() %>%
    addTiles() %>%
    addCircles(radius = ~n)
  
})

```



Column {data-width=300}
-----------------------------------------------------------------------

### Number of Terrorist Incidents

```{r}
renderValueBox({
  valueBox(prettyNum(gtd_filter_af() %>%
                       nrow(), big.mark = ','))
})
```

### Casualties

```{r}
renderPlotly({
  
  plot_ly(gtd_filter_af(), x = ~iyear, y = ~nkill, text = ~gname, type = 'scatter', mode = 'markers',
        marker = list(size = ~nkill, opacity = 0.5)) %>%
  layout(title = 'Number of Casualties',
         xaxis = list(title = "Year", showgrid = FALSE),
         yaxis = list(title = "Number of Casualties", showgrid = FALSE))
})

```

Australasia & Oceania
=========================================================


Column {data-width=250 .sidebar}
-----------------------------------------------------------------------

### Global Terrorist Incidents By Region and Year

```{r}

gtdtib_ao <- filter(gtdtib, region_txt == "Australasia & Oceania")


pickerInput("countryao", 
            "Select a Country In Australasia & Oceania",
            choices = levels(factor(gtdtib_ao$country_txt)),
            selected = levels(factor(gtdtib_ao$country_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

sliderInput("yearao", 
            label = "Select a Year Range:",
            min = 1970,
            max = 1995,
            value = c(1970, 1995),
            step = 1,
            dragRange = TRUE)

pickerInput("attacktypeao", 
            "Select a Type of Terrorist Attack",
            choices = levels(factor(gtdtib_ao$attacktype1_txt)),
            selected = levels(factor(gtdtib_ao$attacktype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

pickerInput("targettypeao", 
            "Select Target Type",
            choices = levels(factor(gtdtib_ao$targtype1_txt)),
            selected = levels(factor(gtdtib_ao$targtype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

gtd_filter_ao <- reactive({
  
  gtdtib_ao %>%
    filter(country_txt %in% input$countryao,
           iyear >= input$yearao[1] & iyear <= input$yearao[2],
           attacktype1_txt %in% input$attacktypeao, 
           targtype1_txt %in% input$targettypeao)
    })

```

Column {data-width=450}
-----------------------------------------------------------------------

### Map of Incidents Around the World

```{r}
renderLeaflet({
  
  gtd_filter_ao() %>%
    group_by(eventid, latitude, longitude) %>%
    count() %>%
    leaflet() %>%
    addTiles() %>%
    addCircles(radius = ~n)
  
})

```



Column {data-width=300}
-----------------------------------------------------------------------

### Number of Terrorist Incidents

```{r}
renderValueBox({
  valueBox(prettyNum(gtd_filter_ao() %>%
                       nrow(), big.mark = ','))
})
```

### Casualties

```{r}
renderPlotly({
  
  plot_ly(gtd_filter_ao(), x = ~iyear, y = ~nkill, text = ~country_txt, type = 'scatter', mode = 'markers',
        marker = list(size = ~nkill, opacity = 0.5)) %>%
  layout(title = 'Number of Casualties',
         xaxis = list(title = "Year", showgrid = FALSE),
         yaxis = list(title = "Number of Casualties", showgrid = FALSE))
})

```

Asia
=========================================================


Column {data-width=250 .sidebar}
-----------------------------------------------------------------------

### Global Terrorist Incidents By Region and Year

```{r}

gtdtib_as <- filter(gtdtib, region_txt %in% c("Central Asia", "South Asia", "East Asia", "Southeast Asia"))


pickerInput("countryas", 
            "Select a Country In Asia",
            choices = levels(factor(gtdtib_as$country_txt)),
            selected = levels(factor(gtdtib_as$country_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

sliderInput("yearas", 
            label = "Select a Year Range:",
            min = 1970,
            max = 1995,
            value = c(1970, 1995),
            step = 1,
            dragRange = TRUE)

pickerInput("attacktypeas", 
            "Select a Type of Terrorist Attack",
            choices = levels(factor(gtdtib_as$attacktype1_txt)),
            selected = levels(factor(gtdtib_as$attacktype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

pickerInput("targettypeas", 
            "Select Target Type",
            choices = levels(factor(gtdtib_as$targtype1_txt)),
            selected = levels(factor(gtdtib_as$targtype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

gtd_filter_as <- reactive({
  
  gtdtib_as %>%
    filter(country_txt %in% input$countryas,
           iyear >= input$yearas[1] & iyear <= input$yearas[2],
           attacktype1_txt %in% input$attacktypeas, 
           targtype1_txt %in% input$targettypeas)
    })

```


Column {data-width=450}
-----------------------------------------------------------------------

### Map of Incidents Around the World

```{r}
renderLeaflet({
  
  gtd_filter_as() %>%
    group_by(eventid, latitude, longitude) %>%
    count() %>%
    leaflet() %>%
    addTiles() %>%
    addCircles(radius = ~n)
  
})

```



Column {data-width=300}
-----------------------------------------------------------------------

### Number of Terrorist Incidents

```{r}
renderValueBox({
  valueBox(prettyNum(gtd_filter_as() %>%
                       nrow(), big.mark = ','))
})
```

### Casualties

```{r}
renderPlotly({
  
  plot_ly(gtd_filter_as(), x = ~iyear, y = ~nkill, text = ~country_txt, type = 'scatter', mode = 'markers',
        marker = list(size = ~nkill, opacity = 0.5)) %>%
  layout(title = 'Number of Casualties',
         xaxis = list(title = "Year", showgrid = FALSE),
         yaxis = list(title = "Number of Casualties", showgrid = FALSE))
})

```

Europe
=========================================================


Column {data-width=250 .sidebar}
-----------------------------------------------------------------------

### Global Terrorist Incidents By Region and Year

```{r}

gtdtib_eu <- filter(gtdtib, region_txt %in% c("Eastern Europe", "Western Europe"))


pickerInput("countryeu", 
            "Select a Country In Europe",
            choices = levels(factor(gtdtib_eu$country_txt)),
            selected = levels(factor(gtdtib_eu$country_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

sliderInput("yeareu", 
            label = "Select a Year Range:",
            min = 1970,
            max = 1995,
            value = c(1970, 1995),
            step = 1,
            dragRange = TRUE)

pickerInput("attacktypeeu", 
            "Select a Type of Terrorist Attack",
            choices = levels(factor(gtdtib_eu$attacktype1_txt)),
            selected = levels(factor(gtdtib_eu$attacktype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

pickerInput("targettypeeu", 
            "Select Target Type",
            choices = levels(factor(gtdtib_eu$targtype1_txt)),
            selected = levels(factor(gtdtib_eu$targtype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

gtd_filter_eu <- reactive({
  
  gtdtib_eu %>%
    filter(country_txt %in% input$countryeu,
           iyear >= input$yeareu[1] & iyear <= input$yeareu[2],
           attacktype1_txt %in% input$attacktypeeu, 
           targtype1_txt %in% input$targettypeeu)
    })

```


Column {data-width=450}
-----------------------------------------------------------------------

### Map of Incidents Around the World

```{r}
renderLeaflet({
  
  gtd_filter_eu() %>%
    group_by(eventid, latitude, longitude) %>%
    count() %>%
    leaflet() %>%
    addTiles() %>%
    addCircles(radius = ~n)
  
})

```



Column {data-width=300}
-----------------------------------------------------------------------

### Number of Terrorist Incidents

```{r}
renderValueBox({
  valueBox(prettyNum(gtd_filter_eu() %>%
                       nrow(), big.mark = ','))
})
```

### Casualties

```{r}
renderPlotly({
  
  plot_ly(gtd_filter_eu(), x = ~iyear, y = ~nkill, text = ~country_txt, type = 'scatter', mode = 'markers',
        marker = list(size = ~nkill, opacity = 0.5)) %>%
  layout(title = 'Number of Casualties',
         xaxis = list(title = "Year", showgrid = FALSE),
         yaxis = list(title = "Number of Casualties", showgrid = FALSE))
})

```

Latin America {data-navmenu="Americas"}
=========================================================


Column {data-width=250 .sidebar}
-----------------------------------------------------------------------

### Global Terrorist Incidents By Region and Year

```{r}

gtdtib_la <- filter(gtdtib, region_txt %in% c("South America", "Central America & Caribbean"))


pickerInput("countryla", 
            "Select a Country In Latin America",
            choices = levels(factor(gtdtib_la$country_txt)),
            selected = levels(factor(gtdtib_la$country_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

sliderInput("yearla", 
            label = "Select a Year Range:",
            min = 1970,
            max = 1995,
            value = c(1970, 1995),
            step = 1,
            dragRange = TRUE)

pickerInput("attacktypela", 
            "Select a Type of Terrorist Attack",
            choices = levels(factor(gtdtib_la$attacktype1_txt)),
            selected = levels(factor(gtdtib_la$attacktype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

pickerInput("targettypela", 
            "Select Target Type",
            choices = levels(factor(gtdtib_la$targtype1_txt)),
            selected = levels(factor(gtdtib_la$targtype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

gtd_filter_la <- reactive({
  
  gtdtib_la %>%
    filter(country_txt %in% input$countryla,
           iyear >= input$yearla[1] & iyear <= input$yearla[2],
           attacktype1_txt %in% input$attacktypela, 
           targtype1_txt %in% input$targettypela)
    })

```


Column {data-width=450}
-----------------------------------------------------------------------

### Map of Incidents Around the World

```{r}
renderLeaflet({
  
  gtd_filter_la() %>%
    group_by(eventid, latitude, longitude) %>%
    count() %>%
    leaflet() %>%
    addTiles() %>%
    addCircles(radius = ~n)
  
})

```



Column {data-width=300}
-----------------------------------------------------------------------

### Number of Terrorist Incidents

```{r}
renderValueBox({
  valueBox(prettyNum(gtd_filter_la() %>%
                       nrow(), big.mark = ','))
})
```

### Casualties

```{r}
renderPlotly({
  
  plot_ly(gtd_filter_la(), x = ~iyear, y = ~nkill, text = ~country_txt, type = 'scatter', mode = 'markers',
        marker = list(size = ~nkill, opacity = 0.5)) %>%
  layout(title = 'Number of Casualties',
         xaxis = list(title = "Year", showgrid = FALSE),
         yaxis = list(title = "Number of Casualties", showgrid = FALSE))
})

```

Middle East & North Africa {data-navmenu="Middle East and Africa"}
===========================================================================


Column {data-width=250 .sidebar}
-----------------------------------------------------------------------

### Global Terrorist Incidents By Region and Year

```{r}

gtdtib_me <- filter(gtdtib, region_txt == "Middle East & North Africa")


pickerInput("countryme", 
            "Select a Country In the Middle East or North Africa",
            choices = levels(factor(gtdtib_me$country_txt)),
            selected = levels(factor(gtdtib_me$country_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

sliderInput("yearme", 
            label = "Select a Year Range:",
            min = 1970,
            max = 1995,
            value = c(1970, 1995),
            step = 1,
            dragRange = TRUE)

pickerInput("attacktypeme", 
            "Select a Type of Terrorist Attack",
            choices = levels(factor(gtdtib_me$attacktype1_txt)),
            selected = levels(factor(gtdtib_me$attacktype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

pickerInput("targettypeme", 
            "Select Target Type",
            choices = levels(factor(gtdtib_me$targtype1_txt)),
            selected = levels(factor(gtdtib_me$targtype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

gtd_filter_me <- reactive({
  
  gtdtib_me %>%
    filter(country_txt %in% input$countryme,
           iyear >= input$yearme[1] & iyear <= input$yearme[2],
           attacktype1_txt %in% input$attacktypeme, 
           targtype1_txt %in% input$targettypeme)
    })

```


Column {data-width=450}
-----------------------------------------------------------------------

### Map of Incidents Around the World

```{r}
renderLeaflet({
  
  gtd_filter_me() %>%
    group_by(eventid, latitude, longitude) %>%
    count() %>%
    leaflet() %>%
    addTiles() %>%
    addCircles(radius = ~n)
  
})

```



Column {data-width=300}
-----------------------------------------------------------------------

### Number of Terrorist Incidents

```{r}
renderValueBox({
  valueBox(prettyNum(gtd_filter_me() %>%
                       nrow(), big.mark = ','))
})
```

### Number of Casualties

```{r}
renderPlotly({
  
  plot_ly(gtd_filter_me(), x = ~iyear, y = ~nkill, text = ~country_txt, type = 'scatter', mode = 'markers',
        marker = list(size = 7, opacity = 0.5)) %>%
 layout(title = 'Number of Casualties',
         xaxis = list(title = "Year", showgrid = FALSE),
         yaxis = list(title = "Number of Casualties", showgrid = FALSE))
})

```



```

North America {data-navmenu="Americas"}
=========================================================


Column {data-width=250 .sidebar}
-----------------------------------------------------------------------

### Global Terrorist Incidents By Region and Year

```{r}

gtdtib_na <- filter(gtdtib, region_txt == "North America")


pickerInput("countryna", 
            "Select a Country In North America",
            choices = levels(factor(gtdtib_na$country_txt)),
            selected = levels(factor(gtdtib_na$country_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

sliderInput("yearna", 
            label = "Select a Year Range:",
            min = 1970,
            max = 1995,
            value = c(1970, 1995),
            step = 1,
            dragRange = TRUE)

pickerInput("attacktypena", 
            "Select a Type of Terrorist Attack",
            choices = levels(factor(gtdtib_na$attacktype1_txt)),
            selected = levels(factor(gtdtib_na$attacktype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

pickerInput("targettypena", 
            "Select Target Type",
            choices = levels(factor(gtdtib_na$targtype1_txt)),
            selected = levels(factor(gtdtib_na$targtype1_txt)),
                             options = pickerOptions(actionsBox = TRUE),
            multiple = TRUE
)

gtd_filter_na <- reactive({
  
  gtdtib_na %>%
    filter(country_txt %in% input$countryna,
           iyear >= input$yearna[1] & iyear <= input$yearna[2],
           attacktype1_txt %in% input$attacktypena, 
           targtype1_txt %in% input$targettypena)
    })

```


Column {data-width=450}
-----------------------------------------------------------------------

### Map of Incidents Around the World

```{r}
renderLeaflet({
  
  gtd_filter_na() %>%
    group_by(eventid, latitude, longitude) %>%
    count() %>%
    leaflet() %>%
    addTiles() %>%
    addCircles(radius = ~n) 
    
  
  
})

```



Column {data-width=300}
-----------------------------------------------------------------------

### Number of Terrorist Incidents

```{r}
renderValueBox({
  valueBox(prettyNum(gtd_filter_na() %>%
                       nrow(), big.mark = ','))
})
```

### Casualties

```{r}
renderPlotly({
  
  plot_ly(gtd_filter_na(), x = ~iyear, y = ~nkill, text = ~country_txt, type = 'scatter', mode = 'markers',
        marker = list(size = ~nkill, opacity = 0.5)) %>%
  layout(title = 'Number of Casualties',
         xaxis = list(title = "Year", showgrid = FALSE),
         yaxis = list(title = "Number of Casualties", showgrid = FALSE))
})

```